---
title: "Practice Coursework"
output:
  html_document: default
---

## Load the Data

```{r}
data<-read.csv('Surgery_20594616.csv')
print(data)
```
## Exploratory Data Analysis

```{r}
# Summary for Current Surgeries
summary(data$Current.surgery)

# Summary for New Surgeries
summary(data$New.surgery)

c_mu <- mean(data$Current.surgery)
c_sd <- sd(data$Current.surgery)
n_mu <- mean(data$New.surgery)
n_sd <- sd(data$New.surgery)
```
### Visualisation
```{r}
boxplot(data$Current.surgery, main = 'Current Surgeries')
boxplot(data$New.surgery, main = 'New Surgeries')
```


```{r}
hist(data$Current.surgery, freq = FALSE, main = 'Histogram for Current Surgeries', xlab = 'Surgery time')
curve(dnorm(x,c_mu,c_sd),3.5,10.9,col=2,add=TRUE)

hist(data$New.surgery, freq = FALSE, main = 'Histogram for New Surgeries', xlab = 'Surgery time')
curve(dnorm(x,n_mu,n_sd),3.5,11.4,col=2,add=TRUE)
```
## Question 1

*Do the data from the new surgical procedure support the hypothesis that the new surgical procedure reduces the mean length of hospital stays?*

```{r}
t.test(data$Current.surgery, data$New.surgery, paired = TRUE, alt = 'greater', var.equal = TRUE)
```
As the p-value is smaller than 0.05 (assuming 5% significance level), we can reject the null 
hypothesis. Therefore, we conclude that the true mean difference is greater than 0 and the new 
surgical procedure has reduced the mean length.

## Question 2

*Using the Matched Case control study data, assess the size of reduction in the length of hospital stay due to the new surgical procedure.*

```{r}
# Calculate the mean difference
mean_difference <- mean(data$Current.surgery - data$New.surgery)

# Construct a confidence interval
conf_interval <- t.test(data$Current.surgery, data$New.surgery, paired = TRUE)$conf.int

# Display the results
mean_difference
conf_interval
```
The mean difference is 0.4103448 and is within the confidence interval.

Lets perform a t-test to confirm that the new procedure has reduced the length of hospital stay by 0.41 days.

```{r}
t.test(data$Current.surgery, data$New.surgery, paired = TRUE, mu = 0.41)
```
Since, the p value is very large, we cannot reject the null hypothesis. Therefore, the true mean 
difference is actually equal to 0.41.

